Multi-resolution Neural Field Estimation

### 2011-03-05 Multiresolution_framework/ltx/manuscript.tex
minor edit
correction to lambdatheta
minor changes
added new plots for kernel estimates
minor edits

\author{\parham{P. Aram}, \dean{D.R. Freestone~\IEEEmembership{Graduate Student Member,~IEEE}}, \mike{M. Dewar}, \ken{K.Scerri}, D.B. Grayden~\IEEEmembership{Member,~IEEE,} and V. Kadirkamanathan~\IEEEmembership{Member,~IEEE} % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{D.\ R.\ Freestone and D.\ B.\ Grayden are with the Department
of Electrical and Electronic Engineering, The University of Melbourne, and The Bionic Ear Institute, VIC, Australia}
\thanks{M. Dewar is with the Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\thanks{K. Scerri is with the Department of Systems and Control Engineering, University
of Malta, Msida, MSD 2080, Malta.}}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

To derive the state-space model, we first define the matrices \mike{bold $r$s here and everywhere (or nowhere)?}
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we pre-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate \cut{the convolution and inner product to form} the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{align}
N_{m}\left(r\right)=&\left(N_{m-1}\ast N_{1}\right)\left(r\right)\nonumber \\
=&\int_0^{1} N_{m-1}\left( r-r'\right)dr' \quad m>1,
\label{SplineConvolutionIntegral}
\end{align}
where $\ast$ denotes convolution and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself since
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline, such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as $N_m\left(r\right)$ is isotropic. The $m^{th}$-order B-spline wavelet function is defined as (see \cite{Chui1992}) 
\begin{align}
 \varphi & _{m}\left(r\right) = \sum_{n=0}^{3m-2} q_n N_{m}\left(2r-n\right) \\
 q & _n = \frac{\left(-1\right)^n}{2^{m-1}} \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$},
\end{align}
where $q_n$ are the coefficients. By exploiting \cut{properties} Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed analytically. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.

% \cut{ It is important to note that B-spline scaling and wavelet functions have the following orthogonality properties \cite{Unser1993}: 
% \begin{equation}
%  \left\langle \psi_{j_1,l_1}(s),\psi_{j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
% \label{PsiPsiOrthogonality}
% \end{equation}
% \begin{equation}
%  \left\langle \phi_{j_1,l_1}(s),\psi_{j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
% \label{PhiPsiOrthogonality}
% \end{equation}
\section{State and Parameter Estimation}
The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Roweis1999,Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. For a multivariate linear dynamic system, the bound used is a quadratic of the form
\begin{equation}\label{eq:Qfunction}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Eq.~\ref{eq:Qfunction} is maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0 & =\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1 & =\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T}\odot \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The operator $\odot$ denotes block-wise matrix multiplication \mike{this still doesn't work! block matrix multiplication isn't a different thing to regular matrix multiplication I think? Isn't it just a different way to do regular matrix multiplication? Also one still doesn't know if the indices in eqs 31-33 are block indices or element indices. Also it's hard to invent notation, and if you do you need to make sure it's defined! Maybe check out the tensor literature? (or simply just index all this properly)}and  $[.]_{i,j}$ denotes $\left(i,j\right)^{th}$ (block) element of the (block) matrix. Each $1\times n_{\theta}$ block of the $n_x \times n_x$ block matrix,  $\mathbf U$ is given by
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix. Note that to separate $\boldsymbol\theta$ from $\boldsymbol\Lambda_{\theta}$ in forming the $\mathcal{Q}$-function we used the relation
\begin{equation}
\left[ \boldsymbol\Lambda_{\theta}\right] _{i,j} =\left[ \mathbf U\right] _{i,j}\boldsymbol\theta^\top
\end{equation}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using Eqs.~\ref{eq:DiscreteTimeModel} and \ref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 100 realizations of 1 second of data were generated. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The distance between adjacent sensors was $0.5$~mm resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed. The spatial cutoff frequency of the observed field, $y_t(r_n)$, was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of \dean{6?}, the cutoff spatial frequency was \dean{$\nu_{cy} = ??$} cycles/mm \cite{Freestone2011}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = ??$ states. \dean{With this number of observations, the estimation problem was well-posed, since the wavelets are orthogonal and the number of states, $n_x < n_y$ at each level of the decomposition.}  
% \cut{Note that wavelets due to their bandpass filter characteristics extract successively higher and higher frequency components by increasing $j$.}

@@

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 2.10$. (b) $j=1$, $n_x=33$, $RMSE = 1.42$. (c) $j=2$, $n_x=65$, $RMSE = 0.80$. (d) $j=3$, $n_x=131$, $RMSE = 0.72$. (e) $j=4$, $n_x=263$, $RMSE=0.72$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/test.pdf}
		\caption{Connectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Weighted kernel scaling functions at $j=1$. (c) Weighted kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

>>>>

\author{\parham{P. Aram}, \dean{D.R. Freestone~\IEEEmembership{Graduate Student Member,~IEEE}}, \mike{M. Dewar}, \ken{K.Scerri~\IEEEmembership{Member,~IEEE,}}, D.B. Grayden~\IEEEmembership{Member,~IEEE,} and V. Kadirkamanathan~\IEEEmembership{Member,~IEEE} % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{D.\ R.\ Freestone and D.\ B.\ Grayden are with the Department
of Electrical and Electronic Engineering, The University of Melbourne, and The Bionic Ear Institute, VIC, Australia}
\thanks{M. Dewar is with the Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\thanks{K. Scerri is with the Department of Systems and Control Engineering, University
of Malta, Msida, MSD 2080, Malta.}}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right) 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we pre-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I,
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate the convolution and inner product to form the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{align}
N_{m}\left(r\right)=&\left(N_{m-1}\ast N_{1}\right)\left(r\right)\nonumber \\
=&\int_0^{1} N_{m-1}\left( r-r'\right)dr' \quad m>1,
\label{SplineConvolutionIntegral}
\end{align}
where $\ast$ denotes convolution and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself since
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline, such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as $N_m\left(r\right)$ is isotropic. The $m^{th}$-order B-spline wavelet function is defined as (see \cite{Chui1992}) 
\begin{align}
 \varphi & _{m}\left(r\right) = \sum_{n=0}^{3m-2} q_n N_{m}\left(2r-n\right) \\
 q & _n = \frac{\left(-1\right)^n}{2^{m-1}} \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$},
\end{align}
where $q_n$ are the coefficients. By exploiting \cut{properties} Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed analytically. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.

% \cut{ It is important to note that B-spline scaling and wavelet functions have the following orthogonality properties \cite{Unser1993}: 
% \begin{equation}
%  \left\langle \psi_{j_1,l_1}(s),\psi_{j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
% \label{PsiPsiOrthogonality}
% \end{equation}
% \begin{equation}
%  \left\langle \phi_{j_1,l_1}(s),\psi_{j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
% \label{PhiPsiOrthogonality}
% \end{equation}
\section{State and Parameter Estimation}
The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Roweis1999,Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. For a multivariate linear dynamic system, the bound used is a quadratic of the form
\begin{equation}\label{eq:Qfunction}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Eq.~\ref{eq:Qfunction} is maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0 & =\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1 & =\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T}\odot \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\odot\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The operator $\odot$ denotes block-wise matrix multiplication \mike{this still doesn't work! block matrix multiplication isn't a different thing to regular matrix multiplication I think? Isn't it just a different way to do regular matrix multiplication? Also one still doesn't know if the indices in eqs 31-33 are block indices or element indices. Also it's hard to invent notation, and if you do you need to make sure it's defined! Maybe check out the tensor literature? (or simply just index all this properly)}and  $[.]_{i,j}$ denotes $\left(i,j\right)^{th}$ (block) element of the (block) matrix. Each $1\times n_{\theta}$ block of the $n_x \times n_x$ block matrix,  $\mathbf U$ is given by
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix. Note that to separate $\boldsymbol\theta$ from $\boldsymbol\Lambda_{\theta}$ in forming the $\mathcal{Q}$-function we used the relation
\begin{equation}
\left[ \boldsymbol\Lambda_{\theta}\right] _{i,j} =\left[ \mathbf U\right] _{i,j}\boldsymbol\theta
\end{equation}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using Eqs.~\ref{eq:DiscreteTimeModel} and \ref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 100 realizations of 1 second of data were generated. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The firing rate slop, $\varsigma$ was set to $0.56  ~mV^{-1}$. The distance between adjacent sensors was $0.5$~mm resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed. The spatial cutoff frequency of the observed field, $y_t(r_n)$, was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of \dean{6?}, the cutoff spatial frequency was \dean{$\nu_{cy} = ??$} cycles/mm \cite{Freestone2011}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = ??$ states. \dean{With this number of observations, the estimation problem was well-posed, since the wavelets are orthogonal and the number of states, $n_x < n_y$ at each level of the decomposition.}  
% \cut{Note that wavelets due to their bandpass filter characteristics extract successively higher and higher frequency components by increasing $j$.}

@@

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 2.10$. (b) $j=1$, $n_x=33$, $RMSE = 1.42$. (c) $j=2$, $n_x=65$, $RMSE = 0.80$. (d) $j=3$, $n_x=131$, $RMSE = 0.72$. (e) $j=4$, $n_x=263$, $RMSE=0.72$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/KernelUnknownSupportWithVarsigma.pdf}
		\caption{Connectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Kernel scaling functions at $j=1$. (c) Kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.
